---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include=FALSE}
# ------------------------------------------------------------------------------
# Knitr defaults for README builds
# ------------------------------------------------------------------------------
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.path  = "man/figures/README-",
  out.width = "100%"
)
```

# ICERglossary

<!-- badges: start -->
<!-- badges: end -->

`ICERglossary` provides a centralized, JSON-based glossary for multilingual
(DE, FR, IT) Large Scale Assessment (LSA) reporting workflows.  
It is designed as a shared backend for Quarto-based reports and R packages
such as `ICERreport` and `ICERplot`, but can also be used standalone as a
lookup and formatting tool. The glossary is intended to be continuously extended 
and reused across different large-scale assessments (LSA), allowing new studies 
and reporting contexts to build on a shared, evolving set of definitions and 
conventions.

```{r, include=FALSE}
# ------------------------------------------------------------------------------
# Load installed package if available; otherwise load from source
# ------------------------------------------------------------------------------
if (!requireNamespace("ICERglossary", quietly = TRUE)) {
  pkgload::load_all(export_all = FALSE, helpers = FALSE, attach_testthat = FALSE)
}
library(ICERglossary)
```

## Installation

```r
devtools::install_git(
  "https://gitlab.switch.ch/unibe-icer/icer-r-packages/icerglossary.git"
)
```

## Key ideas

- **Package-based structure and versioning**
  Maintaining the glossary as an R package enforces a clear structure, 
  proper documentation, and explicit versioning, supporting stable and 
  reproducible reporting workflows.

- **Single source of truth**  
  All central terms, labels, symbols, units, and formatting rules are defined 
  once in JSON and reused consistently across reports and packages, ensuring 
  uniform wording and reducing the risk of errors.

- **Language-aware formatting**  
  Output automatically adapts to DE / FR / IT conventions (wording,
  spacing, symbols).

- **Collaborative development**
  Project contributors can propose new definitions, labels, or conventions via 
  merge requests, enabling shared ownership and gradual refinement of the glossary.

- **Quarto-first, but not Quarto-only**  
  Designed for Quarto inline usage, but fully usable interactively or from
  other R packages.

- **Override-friendly**  
  Project-specific JSON files can temporarily override built-in definitions
  without modifying the package.

## JSON structure (conceptual)

The glossary is currently organized based on top-level keys such as the following
(the corresponding functions are explained in the next section):

- `canton/*`  – Swiss cantons (keys represent the abbreviations used so far)
- `country/*` – PISA countries (keys represent country codes)
- `domain/*` – Various terms used in connection with tested domains
- `feature/*`  – Various terms, value labels etc. used in connection with student characteristics
- `stats/*`  – Statistical terms
- `term/*`  – General terms
- `tech/*`  – Mainly formatting and grammar
- `old/*`  – Old definitions with top-level keys *write* and *draft*


At load time, all relevant JSON files are merged into a single active glossary
object.

## Quick start

### Retrieve labels and switch language

```{r, eval = TRUE, message = TRUE}
options(ICERglossary.verbose = TRUE)

# get canton BE
glcant("BE_d")

# get country France
glcnt("FRA")

# switch language to French
gllang("FR")

# get domain reading (school language)
gldom("sl")

# get feature expression by migration status
glfeat("byMigrationStatus")

# switch language to Italian
gllang("IT")

# get term for sample size
glstat("sample_size")

# get formatting code for points
glformat("points")

glreset()
```

`base_lang` defaults to `"DE"` and acts as a fallback if a key is missing in the
selected language.

### Format numeric values

```{r, eval = TRUE, message = TRUE}
glfmt("points", 125)

glfmt("percentage", 62)

gllang("FR")

glfmt("percentage", 62)

glreset()
```


## Searching and discovery

### Search glossary entries

```{r, eval = TRUE, message = TRUE}
glsearch("sample")
```

### Interactive picker

```{r, eval = FALSE, message = FALSE}
glpick("sample")
```

The picker shows a numbered list of matches and prints a ready-to-paste
function call.

### Quarto picker helper

```{r, eval = FALSE, message = FALSE}
qtpick("sample")
```

Returns function call that can be copied and pasted into Quarto. Example: Pick 
number 2 returns a quarto snippet for sampling size.

```{r, eval = FALSE}
`r {{glstat("sample_size")}}`
```

## Overriding glossary content

Project-specific adjustments can be applied via external JSON files.

### Apply an overlay

```r
"glread("path/to/custom/json")"
```

All `.json` files in the directory are read and merged; existing keys are
overwritten, new keys are added.

### Reset to package defaults

```r
glreset()
```

### Verbose mode

Messages can be enabled or disabled globally:

```r
options(ICERglossary.verbose = TRUE)
```

or per call:

```r
glread("path/to/custom/json", verbose = TRUE)
```

## File layout

- `R/` – package code (API, search helpers, state management)
- `inst/extdata/json/` – built-in glossary JSON files
- `man/` – documentation

## To do

- [x] Core glossary loading and language switching
- [x] Format and symbol resolution
- [x] Interactive search and picker helpers
- [x] Review and stabilize top-level key conventions
- [ ] Align downstream packages (`ICERreport`, `ICERplot`)
- [ ] Think about options to freeze glossary for finished projects

```{r}
packageVersion("ICERglossary")
```

## License

Internal use only (University of Bern / ICER).


